Embedding Intuitionistic into Classical Logic

### 2022-10-07 FoSSaCS/fossacs-EIICL.tex
typos and corrected examples

\begin{lemma}\label{lemma:propositional-normal-form}
	For every propositional formula $\varphi$ there exists an atom $P$ as well as sets of clauses $\mathcal R$ of the form
	$$(A_1\wedge\dots\wedge A_n)\to (B_1\vee\dots\vee B_n)$$
	called \emph{flat} clauses and $\mathcal X$ of the form
	$$(A\to B)\to C$$
	called \emph{implication} clauses such that $\varphi$ is intuitionistically equivalid to
	$$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\right)\to P$$where $A_i, B_i, A, B, C$ are atomic. This size of $\mathcal R$ and $\mathcal X$ is linear in the size of $\varphi$. In general they will contain additional propositional variables.
\end{lemma}

@@

\begin{lemma}\label{lemma:first-order-normal-form}
	For every predicate formula $\varphi$ there exists a nullary relation symol $P$ as well as sets of clauses $\mathcal R$ of the form
	$$\forall \vec x\left((A_1\wedge\dots\wedge A_n)\to (B_1\vee\dots\vee B_n)\right)$$
	called \emph{flat} clauses and $\mathcal X$ of the form
	$$\forall \vec x\left((A\to B)\to C\right)$$
	called \emph{implication} clauses as well as $\mathcal Q$ of the form
	$$\forall\vec x\left((\forall y A)\to B\right)$$
	called \emph{quantification} clauses such that $\varphi$ is intuitionistically equivalid to
	$$\left(\bigwedge\mathcal R\wedge\bigwedge \mathcal X\wedge\bigwedge\mathcal Q\right)\to P$$where $A_i, B_i, A, B, C$ are atomic. This size of $\mathcal R, \mathcal X, \mathcal Q$ is linear in the size of $\varphi$. In general this formula will contain new function and relation symbols.
\end{lemma}

Since we could not find a proof of this result one is given in the appendix\ref{proof:first-order-normal-form}.

@@

The most obvious approach to embedding intuitionistic into classical logic is to express intuitionistic semantics, i.e. Kripke frames, as a first order theory. For every propositional variable $A$ consider a unary predicate $A$ of the same name where $A(u)$ expresses that $A$ is true at some world $u$. We can then naively encode formulas as follows:

@@

\noindent As all we have done is modeling Kripke frames as a first-order theory we directly obtain:

@@

We now argue that if a counter-model exists for $\varphi^\circ$ then a small counter-model exists. This is ultimately what enables our translations.

\begin{definition}
For $\psi\in\mathcal X$ let $[\psi]$ denote the corresponding clause in $\mathcal X^{(u)S}$. Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^{CH}$.
We say $(A\to B)\to C\in\mathcal X$ is \emph{fulfilled} at $m\in M$ iff $A^I(m)\to B^I(m)$ is false or $C^I(m)$ is true. For $\psi\in\mathcal X$ define $$g_\psi : M\to M, m\mapsto\begin{cases}
		m,&\text{ if $\psi$ is fulfilled at $u$,}\\
		f^I_{\lbrack\psi\rbrack}(m),&\text{ else.}		
	\end{cases}$$. Define a model $\mathcal M_T = (M_T, I_T)$ as follows:
	\begin{itemize}
		\item $M_T$ is the set of sequences without repetition on $\mathcal X$.
		\item Interpret $\preceq$ as the prefix-order.
		\item We set $$f_\psi^{I_T}(\psi_1\dots\psi_n) = \begin{cases}
			\psi_1\dots\psi_n, &\text{if $\psi$ occurs in $[\psi_1]\dots[\psi_n]$,}\\
			\psi_1\dots\psi_n\psi, &\text{else.}			
		\end{cases}$$
		\item For propositional variables $P$, we set $$P^{I_T}\left(\psi_1\dots \psi_n\right) = P^I\left(g_{\psi_n}\left(\dots\left(g_{\psi_1}\left(s^I\right)\right)\dots\right)\right).$$
	\end{itemize}
\end{definition}
\begin{lemma}~\label{thm:prop-countermodel-reduction}
	Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^{CH}$.
	\begin{enumerate}
		\item $\psi$ is fulfilled at $f_{[\psi]}^I(u)$ for all $u\in M, \psi\in\mathcal X$.
		\item If $\psi$ is fulfilled at some $m\in M$ then $\psi$ is fulfilled at all $n\succeq m$.
	\end{enumerate}
\end{lemma}

A proof can be found in the appendix as Lemma~\ref{proof:prop-countermodel-reduction}. Essentially what this tells us is that for every $m$ at which $\psi$ is fulfilled, in particular for each $m$ in the image of $f_\psi$, we may add the assumption that $f_\psi(n) = n$ for all $n\geq m$ without changing the validity of $\psi$.
With this lemma in hand showing that $(M_T, I_T)$ is a counter-model to $\varphi^\circ$ is a straightforward check of definitions.

@@

We can apply the previous results to give a direct embedding from intuitionistic into classical propositional logic. We know that $\varphi^{CH}$ is valid iff it is valid for structures that have as a domain the sequences without repetition over $\mathcal F_\to$, ordered by the prefix-relation and $$s^I = \epsilon, f_\psi^I(x) = \begin{cases}
	x, &\text{ if $\psi$ occurs in $x$,}\\
	x\psi, &\text{ otherwise.}
\end{cases}$$
Now we can replace every $\forall$-quantifier by enumerating all the ground terms, and then replace all distinct ground instances of relations by new propositional variables yielding a propositional formula. This gives us the translation stated in Theorem~\ref{thm:reduction-propositional}.
%\begin{theorem}
%	$\bigwedge\mathcal S\to P$ is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset$ is classically valid.
%\end{theorem}

\begin{example}
	Let us consider the law of excluded middle $\varphi = A\vee\neg A$. Then we get
	$$\mathcal S^-(\varphi) = \{(P_A\vee P_{\neg A})\to P_\varphi, A\to P_A, (A\to \bot)\to P_{\neg A}\},$$
	and we know that $\varphi$ is valid iff $\bigwedge \mathcal S^-(\varphi)\to P_\varphi$ is valid. Applying the above lemma this is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset_\varphi$ is classically valid, where, setting $\psi = (A\to \bot)\to P_{\neg A}$,
	\begin{align*}
		\mathcal S^\# =&\{P_\varphi^\epsilon\to P_\varphi^{\psi}, P_A^\epsilon\to P_A^{\psi},P_{\neg A}^\epsilon\to P_{\neg A}^{\psi},A^\epsilon\to A^{\psi},A^\epsilon\to P_A^\epsilon, P_A^{\psi}\to P_A^{\psi}, \}\cup\\ &\{(P_A^\epsilon\vee P_{\neg A}^\epsilon)\to P_\varphi^\epsilon, (P_A^{\psi}\vee P_{\neg A}^{\psi})\to P_\varphi^{\psi},\}\cup\\&\{(A^{\psi}\to \bot)\to P_{\neg A}^\epsilon, (A^{\psi}\to \bot)\to P_{\neg A}^{\psi}\}.
	\end{align*}
	We now note that $P_\varphi^\epsilon  = P_A^\epsilon = P_{\neg A}^\epsilon = A^\epsilon = 0$, $P_{\varphi}^{\psi} = P_A^{\psi} =  P_{\neg A}^{\psi} = A^{\psi} = 1$ indeed defines a counter-model.
\end{example}

@@

Suppose we are given a formula in normal form $$\left(\bigwedge R\wedge\bigwedge X\right)\to P$$ as in Lemma~\ref{lemma:propositional-normal-form}. We will now formulate a QBF that expresses that $\varphi$ has an intuitionistic counter-model. The idea how to get a polynomially-sized formula is: Instead of expressing validity at every node of $\mathcal M_T$ --- which was done in the CPC translation --- we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model. The universally quantified variables $F_\psi$ in the next definition handle the branching, i.e. express which path is considered.

\begin{definition}
	For every propositional variable $A$ occurring in $\mathcal S$ and non-negative integer $n$ consider a new atom $A^n$ and for every formula $\psi\in\mathcal X$ consider a new atom $X_\psi^n$. Let $\vec A^n$ range over all propositional variables $A^n$ and $\vec X^n$ range over all $X_\psi^n$. Define
	\begin{itemize}
		\item {\normalfont Valid}$(n)$ which encodes that $\vec X^n_\psi$ represents a valid next element of a sequence without repetition, i.e. if we interpret the $\vec X^i$ as bit-vectors, then $\vec X^n$ has exactly one bit set to $1$, indicating the formula at the $n$-th position of the sequence, and it is at a position that is $0$ in $\sum\{\vec X^i\:|\:i < n\}$.
		\item {\normalfont Persistent}$(n)$ which encodes that the persistency condition holds.
		\item {\normalfont Sat}$_{\mathcal R}(n)$ encoding that the formulas in $\mathcal R$ hold at the $n$-th world of the path.
		\item {\normalfont Sat}$_{\mathcal X}(n)$ encoding that the formula in $\mathcal X$ that is represented by $\vec X^{n-1}$ holds.
	\end{itemize}
	Have $k = |\mathcal X|$ and define$$\varphi^Q_{i} = \exists \vec A^i\forall \vec X^i\left(\text{\normalfont Persistent}(i)\wedge \text{\normalfont Sat}_{\mathcal R}(i)\wedge \text{\normalfont Sat}_{\mathcal X}(i)\wedge\left(\text{\normalfont Valid}(i)\to \varphi^Q_{i+1}\right)\right)$$
	for $0 < i < k$, as well as the special cases
	$$\varphi^Q_{k} = \exists \vec A^k\left(\text{\normalfont Persistent}(k)\wedge \text{\normalfont Sat}_{\mathcal R}(k)\wedge \text{\normalfont Sat}_{\mathcal X}(k)\right)$$
	for leafs, and $$\varphi^Q = \varphi_0^Q = \exists \vec A^0\forall \vec X^0\left(\neg P^0\wedge \text{\normalfont Sat}_{\mathcal R}(0)\wedge(\text{\normalfont Valid}(0)\to \varphi_1^Q)\right)$$for the root.
	Example encodings of the above formulas are:
	\begin{align*}
		\text{\normalfont Valid}(n) = &\left(\bigvee \{X^n_\psi\:|\:\psi\in\mathcal X\}\right)\wedge\left(\bigwedge \{\neg(X^n_{\psi_1}\wedge X^n_{\psi_2})\:|\:\psi_1\neq\psi_2\in\mathcal X\}\right)\wedge\\&\left(\bigwedge\left\{\bigwedge\{X^i_\psi\to\neg X^n_\psi\:|\:i < n\}\:|\:\psi\in\mathcal X\right\}\right)\\
		\text{\normalfont Persistent}(n) = & \bigwedge\{A^{n-1}\to A^n\:|\: A \text{ prop. variable with }A^{n-1}\in\vec A^{n-1}, A^n \in\vec A^{n}\}\\
		\text{\normalfont Sat}_{\mathcal R}(n) = &\bigwedge\left\{(A^n\circ B^n)\to C^n\:|\:(A\circ B)\to C\in \mathcal R\right\}\wedge\\
		&\bigwedge\left\{A^n\to (B^n\circ C^n)\:|\:A\to (B\circ C)\in \mathcal R\right\}\\
		\text{\normalfont Sat}_{\mathcal X}(n) = & \bigwedge\{X^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal X\}
	\end{align*}
\end{definition}

\begin{example}
	With the previous encoding for double negation elimination $\varphi = ((A\to \bot)\to \bot)\to A$ we have
	$$\hspace*{-.3cm}
	\varphi^Q = \exists A^0\forall X^0\left(\neg A^0\wedge\left(X^0\to \exists A^1\left((A^0\to A^1)\wedge(X_0\to (A^1\to \bot)\to \bot)\right)\right)\right),
	$$
	which is a satisfiable QBF since $A^0 = 0, A^1 = 1$ satisfies it for any $X^0$.
\end{example}

@@

\begin{lemma}
	For every predicate formula $\varphi$ there exists a nullary relation symol $P$ as well as sets of clauses $\mathcal R$ of the form
	$$\forall \vec x\left((A_1\wedge\dots\wedge A_n)\to (B_1\vee\dots\vee B_n)\right)$$
	called \emph{flat} clauses and $\mathcal X$ of the form
	$$\forall \vec x\left((A\to B)\to C\right)$$
	called \emph{implication} clauses as well as $\mathcal Q$ of the form
	$$\forall\vec x\left((\forall y A)\to B\right)$$
	called \emph{quantification} clauses such that $\varphi$ is intuitionistically equivalid to
	$$\left(\bigwedge\mathcal R\wedge\bigwedge \mathcal X\wedge\bigwedge\mathcal Q\right)\to P$$where $A_i, B_i, A, B, C$ are atomic. This size of $\mathcal R, \mathcal X, \mathcal Q$ is linear in the size of $\varphi$. In general this formula will contain new function symbols and relational symbols.
\end{lemma}

\begin{proof}\label{proof:first-order-normal-form}
	As in~\cite{otten2005clausal} we may assume that $\varphi$ is of the form $\psi\to P$. Recall that the transformation works by transforming $\psi$ into a set of equivalent assumptions $\left\lbrack \psi\right\rbrack$ of the desired form. We extend the given procedure with rules for quantifiers as follows:
	\begin{align*}
		\left\lbrack(\forall x \psi)\to A\right\rbrack&:= \left\lbrack\psi[s(\vec z)/x]\to A\right\rbrack\\
		\left\lbrack A\to (\forall x\psi)\right\rbrack&:= \{\forall x\chi[x/a]\:|\:\chi\in\left\lbrack A\to\psi[a/x]\right\rbrack\}\\
		\left\lbrack(\exists x\psi)\to A\right\rbrack&:= \{\forall x\chi[x/a]\:|\:\chi\in\left\lbrack A\to\psi[a/x]\right\rbrack\}\\
		A\to (\exists x\psi)&:= \left\lbrack A\to\psi[s(\vec z)/x]\right\rbrack
	\end{align*}
	where $s$ is a new function symbol, $a$ is a new free variable and $\vec z$ ranges over the free variables in $\psi$.
	While the generated set of clauses is in general not equivalent to $\psi$, section~\ref{section:herbrandiaztion} guarantees that it is at least equivalid.
\end{proof}

>>>>

\begin{lemma}\label{lemma:propositional-normal-form}
	For every propositional formula $\varphi$ there exists an atom $P$ as well as sets of clauses $\mathcal R, \mathcal X$, with $\mathcal R$ containing  \emph{flat clauses} of the form
	$$(A_1\wedge\dots\wedge A_n)\to (B_1\vee\dots\vee B_n)$$
	and $\mathcal X$ containing \emph{implication clauses} of the form
	$$(A\to B)\to C,$$
	such that $\varphi$ is intuitionistically equivalid to
	$$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\right)\to P$$where $A_i, B_i, A, B, C$ are atomic. This size of $\mathcal R$ and $\mathcal X$ is linear in the size of $\varphi$. In general they will contain additional propositional variables.
\end{lemma}

@@

\begin{lemma}\label{lemma:first-order-normal-form}
	For every predicate formula $\varphi$ there exists a nullary relation symol $P$ as well as sets of clauses $\mathcal R,\mathcal X, \mathcal Q$, with $\mathcal R$ containing \emph{flat clauses} of the form
	$$\forall \vec x\left((A_1\wedge\dots\wedge A_n)\to (B_1\vee\dots\vee B_n)\right),$$
	$\mathcal X$ containing \emph{implication clauses} of the form
	$$\forall \vec x\left((A\to B)\to C\right)$$
	and $\mathcal Q$ containing \emph{quantification clauses} of the form
	$$\forall\vec x\left((\forall y A)\to B\right),$$
	such that $\varphi$ is intuitionistically equivalid to
	$$\left(\bigwedge\mathcal R\wedge\bigwedge \mathcal X\wedge\bigwedge\mathcal Q\right)\to P$$where $A_i, B_i, A, B, C$ are atomic. This size of $\mathcal R, \mathcal X, \mathcal Q$ is linear in the size of $\varphi$. In general this formula will contain new function and relation symbols.
\end{lemma}

Since we could not find a proof of this result one is given in the appendix as Lemma~\ref{proof:first-order-normal-form}.

@@

The most obvious approach to embedding intuitionistic into classical logic is to examine intuitionistic semantics, e.g. Kripke frames, as a classical first order theory. For every propositional variable $A$ consider a unary predicate $A$ of the same name where $A(u)$ expresses that $A$ is true at some world $u$. We can then consider the following naive encoding:

@@

\noindent As all we have done is modeling Kripke frames as a first-order theory so we directly obtain:

@@

We now argue that if a counter-model for $\varphi^{CH}$ exists then a small counter-model exists. This is ultimately what enables our translations.

\begin{definition}
For $\psi\in\mathcal X$ let $[\psi]$ denote the corresponding clause in $\mathcal X^{(u)S}$. Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^{CH}$.
We say $(A\to B)\to C\in\mathcal X$ is \emph{fulfilled} at $m\in M$ iff $A^I(m)\to B^I(m)$ is false or $C^I(m)$ is true. For $\psi\in\mathcal X$ define $$g_\psi : M\to M, m\mapsto\begin{cases}
		m,&\text{ if $\psi$ is fulfilled at $u$,}\\
		f^I_{\lbrack\psi\rbrack}(m),&\text{ else.}		
	\end{cases}$$Define a model $\mathcal M_T = (M_T, I_T)$ as follows:
	\begin{itemize}
		\item $M_T$ is the set of sequences without repetition on $\mathcal X$.
		\item Interpret $\preceq$ as the prefix-order.
		\item We set $$f_\psi^{I_T}(\psi_1\dots\psi_n) = \begin{cases}
			\psi_1\dots\psi_n, &\text{if $\psi$ occurs in $[\psi_1]\dots[\psi_n]$,}\\
			\psi_1\dots\psi_n\psi, &\text{else.}			
		\end{cases}$$
		\item For propositional variables $P$, we set $$P^{I_T}\left(\psi_1\dots \psi_n\right) = P^I\left(g_{\psi_n}\left(\dots\left(g_{\psi_1}\left(s^I\right)\right)\dots\right)\right).$$
	\end{itemize}
\end{definition}
\begin{lemma}~\label{thm:prop-countermodel-reduction}
	Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^{CH}$.
	\begin{enumerate}
		\item $\psi$ is fulfilled at $f_{[\psi]}^I(u)$ for all $u\in M, \psi\in\mathcal X$.
		\item If $\psi$ is fulfilled at some $m\in M$ then $\psi$ is fulfilled at all $n\succeq m$.
	\end{enumerate}
\end{lemma}

A proof can be found in the appendix as Lemma~\ref{proof:prop-countermodel-reduction}. Essentially what this tells us is that for every $m$ at which $\psi$ is fulfilled, in particular for each $m$ in the image of $f_{[\psi]}$, we may add the assumption that $f_{[\psi]}(n) = n$ for all $n\geq m$ without changing the validity of $\psi$.
With this lemma in hand showing that $(M_T, I_T)$ is a counter-model to $\varphi^{CH}$ is a straightforward check of definitions.

@@

We can apply the previous results to give a direct embedding from intuitionistic into classical propositional logic. We know that $\varphi^{CH}$ is valid iff it is valid for structures that have as a domain the sequences without repetition over $\mathcal F_\to$, ordered by the prefix-relation and $$s^I = \epsilon, f_\psi^I(x) = \begin{cases}
	x, &\text{ if $\psi$ occurs in $x$,}\\
	x\psi, &\text{ otherwise.}
\end{cases}$$
Now we can replace every $\forall$-quantifier in $\varphi^{CH}$ by enumerating all the ground terms, and then replace all distinct ground instances of relations by new propositional variables yielding a propositional formula. Note that the encoding of partial order becomes redundant with this so we drop it, but we need to keep the persistency restraints. This gives us the following theorem:

\begin{theorem}
	Suppose we are given a formula as in Lemma~\ref{lemma:propositional-normal-form}, i.e. of the form
	$$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal S\right)\to P$$
	with flat $\mathcal R$ and implication $\mathcal S$. Let $\Lambda$ denote the set of sequences without repetition over $\mathcal X$. For every propositional variable $A$ and $\lambda\in\Lambda$ consider a new propositional Variable $X^\lambda$. Define $R^\#$ as the union over all sets
	$$
		\{(A_1^\lambda \wedge\dots\wedge A_n^\lambda)\to(B_1^\lambda \vee\dots\vee B_n^\lambda)\:|\:\lambda\in\Lambda\}
	$$
	for $(A_1 \wedge\dots\wedge A_n)\to(B_1\vee\dots\vee B_n)\in\mathcal R$ and $\mathcal X^\#$ as the union over all sets
	$$
		\{({A^\lambda\psi} \to A^{\lambda\psi})\to(C^\lambda)\:|\:\lambda\in\Lambda, \psi\text{ does not occur in }\lambda\}
	$$
	for $\psi = (A\to B)\to C\in\mathcal X$ as well as $P^\#$ as the union over all sets
	$$
		\{A^\lambda\to A^{\lambda\psi}\:|\:\lambda\in\Lambda, \psi\psi\text{ does not occur in }\lambda\}
	$$
	 for each propositional variable $A$ occurring in $\varphi$ encoding persistency. Then
	$$
		\varphi^\# := \left(\bigwedge P^\#\wedge\bigwedge \mathcal R^\#\wedge\bigwedge\mathcal X^\#\right)\to P^\epsilon
	$$
	is classically valid if and only if $\varphi$ is intuitionistically valid.
\end{theorem}
%\begin{theorem}
%	$\bigwedge\mathcal S\to P$ is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset$ is classically valid.
%\end{theorem}

\begin{example}
	Let us consider the law of excluded middle $\varphi = A\vee\neg A$ for which there is the equivalid normal form
	$$((A\to P)\wedge ((A\to \bot)\to P))\to P.$$
	Applying the above lemma this is intuitionistically valid if and only if $$((A^\epsilon\to A^\psi)\wedge(P^\epsilon\to P^\psi)\wedge(A^\epsilon\to P^\epsilon)\wedge(A^\psi\to P^\psi)\wedge ((A^\psi\to \bot)\to P^\epsilon))\to P^\epsilon$$ is classically valid where $\psi := (A\to \bot)\to P$.
	We note that $A^\epsilon  = P^\epsilon = 0$, $A^\psi = P^{\psi}  = 1$ indeed defines a counter-model.
\end{example}

@@

Suppose we are given a formula $\varphi$ in normal form $$\left(\bigwedge R\wedge\bigwedge X\right)\to P$$ as in Lemma~\ref{lemma:propositional-normal-form}. We will now formulate a QBF that expresses that $\varphi$ has an intuitionistic counter-model. The idea how to get a polynomially-sized formula is: Instead of expressing validity at every node of $\mathcal M_T$ --- which was done in the CPC translation --- we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model. The universally quantified variables $X_\psi$ in the next definition handle the branching, i.e. express which path is considered.

\begin{definition}
	For every propositional variable $A$ occurring in $\varphi$ and non-negative integer $n$ consider a new propositional variable $A^n$ and for every formula $\psi\in\mathcal X$ consider a new propositional variable $X_\psi^n$. Let $\vec A^n$ range over all propositional variables $A^n$ and $\vec X^n$ range over all $X_\psi^n$. Define
	\begin{itemize}
		\item {\normalfont Valid}$(n)$ which encodes that $\vec X^n_\psi$ represents a valid next element of a sequence without repetition, i.e. if we interpret the $\vec X^i$ as bit-vectors, then $\vec X^n$ has exactly one bit set to $1$, indicating the formula at the $n$-th position of the sequence, and it is at a position that is $0$ in $\sum\{\vec X^i\:|\:i < n\}$.
		\item {\normalfont Persistent}$(n)$ which encodes that the persistency condition holds.
		\item {\normalfont Sat}$_{\mathcal R}(n)$ encoding that the formulas in $\mathcal R$ hold at the $n$-th world of the path.
		\item {\normalfont Sat}$_{\mathcal X}(n)$ encoding that the formula in $\mathcal X$ that is represented by $\vec X^{n-1}$ holds.
	\end{itemize}
	Have $k = |\mathcal X|$ and define$$\varphi^Q_{i} = \exists \vec A^i\forall \vec X^i\left(\text{\normalfont Persistent}(i)\wedge \text{\normalfont Sat}_{\mathcal R}(i)\wedge \text{\normalfont Sat}_{\mathcal X}(i)\wedge\left(\text{\normalfont Valid}(i)\to \varphi^Q_{i+1}\right)\right)$$
	for $0 < i < k$, as well as the special cases
	$$\varphi^Q_{k} = \exists \vec A^k\left(\text{\normalfont Persistent}(k)\wedge \text{\normalfont Sat}_{\mathcal R}(k)\wedge \text{\normalfont Sat}_{\mathcal X}(k)\right)$$
	for leafs, and $$\varphi^Q = \varphi_0^Q = \exists \vec A^0\forall \vec X^0\left(\neg P^0\wedge \text{\normalfont Sat}_{\mathcal R}(0)\wedge(\text{\normalfont Valid}(0)\to \varphi_1^Q)\right)$$for the root.
	Example encodings of the above formulas are:
	\begin{align*}
		\text{\normalfont Valid}(n) = &\left(\bigvee \{X^n_\psi\:|\:\psi\in\mathcal X\}\right)\wedge\left(\bigwedge \{\neg(X^n_{\psi_1}\wedge X^n_{\psi_2})\:|\:\psi_1\neq\psi_2\in\mathcal X\}\right)\wedge\\&\left(\bigwedge\left\{\bigwedge\{X^i_\psi\to\neg X^n_\psi\:|\:i < n\}\:|\:\psi\in\mathcal X\right\}\right)\\
		\text{\normalfont Persistent}(n) = & \bigwedge\{A^{n-1}\to A^n\:|\: A \text{ prop. variable with }A^{n-1}\in\vec A^{n-1}, A^n \in\vec A^{n}\}\\
		\text{\normalfont Sat}_{\mathcal R}(n) = &\bigwedge\left\{(A^n\circ B^n)\to C^n\:|\:(A\circ B)\to C\in \mathcal R\right\}\wedge\\
		&\bigwedge\left\{A^n\to (B^n\circ C^n)\:|\:A\to (B\circ C)\in \mathcal R\right\}\\
		\text{\normalfont Sat}_{\mathcal X}(n) = & \bigwedge\{X^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal X\}
	\end{align*}
\end{definition}

\begin{example}
	With the previous encoding for double negation elimination $\varphi = ((A\to \bot)\to \bot)\to A$ we have
	$$\hspace*{-.3cm}
	\varphi^Q = \exists A^0\forall X^0\left(\neg A^0\wedge\left(X^0\to \exists A^1\left((A^0\to A^1)\wedge(X_0\to (A^1\to \bot)\to \bot)\right)\right)\right),
	$$
	which is a satisfiable QBF since $A^0 = 0, A^1 = 1$ satisfies it for any choice of $X^0$.
\end{example}

@@

\begin{lemma}\label{proof:first-order-normal-form}
	For every predicate formula $\varphi$ there exists a nullary relation symol $P$ as well as sets of clauses $\mathcal R$ of the form
	$$\forall \vec x\left((A_1\wedge\dots\wedge A_n)\to (B_1\vee\dots\vee B_n)\right)$$
	called \emph{flat} clauses and $\mathcal X$ of the form
	$$\forall \vec x\left((A\to B)\to C\right)$$
	called \emph{implication} clauses as well as $\mathcal Q$ of the form
	$$\forall\vec x\left((\forall y A)\to B\right)$$
	called \emph{quantification} clauses such that $\varphi$ is intuitionistically equivalid to
	$$\left(\bigwedge\mathcal R\wedge\bigwedge \mathcal X\wedge\bigwedge\mathcal Q\right)\to P$$where $A_i, B_i, A, B, C$ are atomic. This size of $\mathcal R, \mathcal X, \mathcal Q$ is linear in the size of $\varphi$. In general this formula will contain new function symbols and relational symbols.
\end{lemma}

\begin{proof}
	As in~\cite{otten2005clausal} we may assume that $\varphi$ is of the form $\psi\to P$. Recall that the transformation works by transforming $\psi$ into a set of equivalent assumptions $\left\lbrack \psi\right\rbrack$ of the desired form. We extend the given procedure with rules for quantifiers as follows:
	\begin{align*}
		\left\lbrack(\forall x \psi)\to A\right\rbrack&:= \left\lbrack\psi[s(\vec z)/x]\to A\right\rbrack\\
		\left\lbrack A\to (\forall x\psi)\right\rbrack&:= \{\forall x\chi[x/a]\:|\:\chi\in\left\lbrack A\to\psi[a/x]\right\rbrack\}\\
		\left\lbrack(\exists x\psi)\to A\right\rbrack&:= \{\forall x\chi[x/a]\:|\:\chi\in\left\lbrack A\to\psi[a/x]\right\rbrack\}\\
		A\to (\exists x\psi)&:= \left\lbrack A\to\psi[s(\vec z)/x]\right\rbrack
	\end{align*}
	where $s$ is a new function symbol, $a$ is a new free variable and $\vec z$ ranges over the free variables in $\psi$.
	While the generated set of clauses is in general not equivalent to $\psi$, section~\ref{section:herbrandiaztion} guarantees that it is at least equivalid.
\end{proof}
